Updated Constraints on Self-Interacting Dark Matter from Supernova 1987A

### 2017-05-14 draft.tex
Wrote conclusion section. Minor edits throughout.

%--- journals
\newcommand{\physrep}{Phys. Reports}
\newcommand{\mnras}{Mon. Not. R. Astron. Soc.}
\newcommand{\apjl}{Astrophys. J. Lett.}
\newcommand{\jcap}{J. Cosmol. Astropart. Phys.}

@@

Having obtained an expression for $ \kappa $, we can now find the optical depth as follows. 
Define a typical optical depth as $\tau_\mathrm{R} = \kappa_\mathrm{R} \rho_\mathrm{R} R $. We then have 
\beq 
\kappa \rho R = \tau_\mathrm{R} \left( \frac{\rho}{\rho_\mathrm{R}} \right)^2 \left( \frac{T_\mathrm{R}}{T} \right)^{3/2}.
\eeq
%
This can be combined with the previous expressions for the density and temperature as a function of position 
and plugged in to the integral expression for the optical depth to obtain 
\bea 
\tau &=& \int_{r_x}^{\infty}\, \tau_\mathrm{R} \left( \frac{R}{r} \right)^{3n/2} \dd r \\
 &=& \frac{\tau_\mathrm{r}}{\frac{3n}{2}-1} \left( \frac{T_\mathrm{A'}}{T_\mathrm{R}} \right)^{(9/2-3/n)}. 
 \eea
 The bound on the coupling is obtained by requiring  $ \tau(\varepsilon,m_\mathrm{A'}) \le 2/3 $. Note that strictly speaking we should have determined $ r_{\mathrm{dp}}$ rather than assuming a value. However, it is possible to verify the self-consistency of our assumption using this model for the supernova atmosphere. In particular, this model implies that an optical depth of $\tau=2/3$ is reached at a radial position of $ r_{\mathrm{dp}}=XXXX$,  validating the assumption made at the outset. \arz{Put in the missing number here.}

As we stated in the previous subsection, the approximate treatment of the upper limit of the exclusion region 
is justified by the fact that values of $\varepsilon$ close to the upper limit of the exclusion region are 
ruled out by independent, terrestrial experiments.

@@

Our constraints are on the mixing parameter are shown in 
Fig.~\ref{fig:constraint} and agree well with the those presented in 
Ref.~\cite{rrapaj_reddy16} and, for dark photon masses $m_{\mathrm{A}'} \gtrsim 10$~MeV, 
Ref.~\cite{hardy_lasenby17}. We argue that the lower bound on the dark photon mixing parameter 
is only sensitive to the SRA to within a factor of $\sim 3$ or less. A more precise constraint 
on such models from SN1987A is probably not practicable because the interior temperature 
of the SN1987A core is somewhat uncertain and this uncertainty alone gives rise to an uncertainty 
in the excluded region of at least a factor of $\sim 2-3$. Finally, \citet{hardy_lasenby17} consider 
plasma effects within the supernova. The most important ramification of this work in the context of 
our paper is that it suggests significantly weaker constraints on $\varepsilon$ in the mass 
range $m_{\mathrm{A}'} \lesssim 10$~MeV.

@@

AKL was supported in part by NSF grant PHY-1519175. 
ARZ was supported in part by the DoE through Grant DE-SC0007914 and by 
the Pittsburgh Particle physics Astrophysics and Cosmology Center (Pitt PACC) 
at the University of Pittsburgh.

>>>>

%--- journals
\newcommand{\physrep}{Phys. Reports}
\newcommand{\mnras}{Mon. Not. R. Astron. Soc.}
\newcommand{\apjl}{Astrophys. J. Lett.}
\newcommand{\apjs}{Astrophys. J. Supp. Ser.}
\newcommand{\jcap}{J. Cosmol. Astropart. Phys.}

@@

Having obtained an expression for $ \kappa $, we can now find the optical depth as follows. 
Define a typical optical depth as $\tau_\mathrm{R} = \kappa_\mathrm{R} \rho_\mathrm{R} R $. We then have 
\beq 
\kappa \rho R = \tau_\mathrm{R} \left( \frac{\rho}{\rho_\mathrm{R}} \right)^2 \left( \frac{T_\mathrm{R}}{T} \right)^{3/2}.
\eeq
%
This can be combined with the previous expressions for the density and temperature as a function of position 
and plugged in to the integral expression for the optical depth to obtain 
\bea 
\tau &=& \int_{r_x}^{\infty}\, \tau_\mathrm{R} \left( \frac{R}{r} \right)^{3n/2} \dd r \\
 &=& \frac{\tau_\mathrm{r}}{\frac{3n}{2}-1} \left( \frac{T_\mathrm{A'}}{T_\mathrm{R}} \right)^{(9/2-3/n)}. 
 \eea
 The bound on the coupling is obtained by requiring  $ \tau(\varepsilon,m_\mathrm{A'}) \le 2/3 $. Note that strictly speaking we should have determined $ r_{\mathrm{dp}}$ rather than assuming a value. However, it is possible to verify the self-consistency of our assumption using this model for the supernova atmosphere. In particular, this model implies that an optical depth of $\tau=2/3$ is reached at a radial position of $ r_{\mathrm{dp}}=XXXX$,  validating the assumption made at the outset. \arz{Put in the missing number here.} As we stated in the previous subsection, the approximate treatment of the upper limit of the exclusion region is justified by the fact that values of $\varepsilon$ close to the upper limit of the exclusion region are ruled out by independent, terrestrial experiments.

@@

Our constraints are on the mixing parameter are shown in 
Fig.~\ref{fig:constraint} and agree well with the those presented in 
Ref.~\cite{rrapaj_reddy16} and, for dark photon masses $m_{\mathrm{A}'} \gtrsim 10$~MeV, 
Ref.~\cite{hardy_lasenby17}. We argue that the lower bound on the dark photon mixing parameter 
is only sensitive to the SRA to within a factor of $\sim 3$ or less. A more precise constraint 
on such models from SN1987A is probably not practicable because the interior temperature 
of the SN1987A core is somewhat uncertain and this uncertainty alone gives rise to an uncertainty 
in the excluded region of at least a factor of $\sim 2-3$. Finally, \citet{hardy_lasenby17} consider 
plasma effects within the supernova. The most important ramification of this work in the context of 
our paper is that it demonstrates that constraints on $\varepsilon$ should be significantly weakened 
relative to our results in the mass range $m_{\mathrm{A}'} \lesssim 10$~MeV.

Finally, we note that the interior temperature and density within the progenitor of SN1987A remain 
uncertain at least at the level of tens of percent \citep[e.g.][]{perego15,farmer16}. This is well worth 
noting because the constraints one derives using the techniques in this work or similar techniques 
are quite sensitive to the properties of the SN1987A core, particularly the core temperature. Though 
we do not show this result explicitly, we have replicated our calculations for shifts in core temperature 
of $\pm 5\, \mathrm{MeV}$ and find that the boundaries of the excluded region shift by a factor of $\sim 2-3$ at low 
mass ($m_{\mathrm{A}'} \lesssim T$ where $T = 30\, \mathrm{MeV}$ in our fiducial calculation, following 
work or previous authors) and considerably more at high mass. 
We conclude that the differences among theoretical techniques for 
computing the excluded region (OPE, EFT, SRA, ...) lead to uncertainties that are of the same order or 
smaller than the uncertainty in the excluded region induced by our limited knowledge of the 
properties of the SN1987A progenitor interior. The excluded regions we and other authors delineate 
should be considered to be uncertain by a factor or $\sim 3$. Refining the theoretical approach to this 
problem is therefore unwarranted until a breakthrough in our understanding of the interior of the 
SN1987A progenitor is achieved.

@@

We thank Carles Badenes, John Hillier, and Francis Timmes for guidance in understanding 
the state of knowledge of the SN1987A progenitor. 
AKL was supported in part by NSF grant PHY-1519175. 
ARZ was supported in part by the DoE through Grant DE-SC0007914 and by 
the Pittsburgh Particle physics Astrophysics and Cosmology Center (Pitt PACC) 
at the University of Pittsburgh.
